---
title: "Common Text Processing Tasks with textworks"
author: "Martin Chan"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Common Text Processing Tasks with textworks}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
# Load the textworks package
library(textworks)
```

## Common Text Processing Workflows

This vignette demonstrates practical applications of textworks functions for common text processing tasks you might encounter in data analysis and visualization.

### Cleaning Survey Responses

```{r survey-cleaning}
# Raw survey responses with various formatting issues
responses <- c(
  "very_satisfied_with_service",
  "somewhatDissatisfied",
  "neutralResponse", 
  "EXTREMELY_SATISFIED",
  "not_sure_about_this"
)

# Clean and standardize the responses
cleaned_responses <- responses %>%
  us_to_space() %>%           # Remove underscores
  camel_clean() %>%           # Fix CamelCase
  sentence_case()             # Standardize capitalization

data.frame(
  Original = responses,
  Cleaned = cleaned_responses
)
```

### Preparing Text for Visualization

```{r viz-prep}
# Long category names that need line breaks for plots
category_names <- c(
  "Customer Service and Support Quality Assessment",
  "Product Development and Innovation Strategy",
  "Marketing and Brand Management Effectiveness"
)

# Add line breaks for better display in charts
short_names <- sapply(category_names, function(x) {
  add_line_breaks(x, nword = 3)
})

cat("Original vs. Formatted:\n")
for(i in 1:length(category_names)) {
  cat("\nOriginal:", category_names[i])
  cat("\nFormatted:\n", short_names[i], "\n")
}
```

### Processing File Names and Paths

```{r file-processing}
# Messy file names that need cleaning
file_names <- c(
  "data_analysis_report_2023.xlsx",
  "customerFeedbackSummary.pdf",
  "QUARTERLY_RESULTS_Q4.pptx"
)

# Create clean display names
display_names <- file_names %>%
  # Remove file extensions for display
  sub("\\.[^.]*$", "", .) %>%
  # Clean formatting
  us_to_space() %>%
  camel_clean() %>%
  cap_first()

data.frame(
  File = file_names,
  Display = display_names
)
```

### Text Analysis and Metrics

```{r text-analysis}
# Sample text for analysis
sample_text <- "The quick brown fox jumps over the lazy dog. This pangram contains every letter of the alphabet at least once."

# Basic text metrics
cat("Text Analysis:\n")
cat("Total words:", nwords(sample_text), "\n")
cat("Total characters:", nchar(sample_text), "\n")
cat("First 20 characters:", str_left(sample_text, 20), "\n")
cat("Last 15 characters:", str_right(sample_text, 15), "\n")

# Analyze word patterns
cat("\nMost common bigrams:\n")
print(count_ngram(sample_text, n = 2))
```

### Creating Formatted Output

```{r formatted-output}
# Data that needs formatting for reports
metrics <- data.frame(
  metric = c("total_users", "monthly_revenue", "customer_satisfaction"),
  value = c(125000, 2500000, 87.5)
)

# Format the metrics nicely
metrics$formatted_metric <- cap_first(us_to_space(metrics$metric))
metrics$formatted_value <- ifelse(
  metrics$metric == "customer_satisfaction",
  paste0(metrics$value, "%"),
  comma(metrics$value)
)

# Display formatted results
cat("Formatted Metrics Report:\n")
for(i in 1:nrow(metrics)) {
  cat(sprintf("%s: %s\n", 
              metrics$formatted_metric[i], 
              metrics$formatted_value[i]))
}
```

This vignette shows how textworks functions can be combined to solve real-world text processing challenges efficiently and cleanly.
