---
title: "Introduction to textworks: Text Wrangling with Wittgenstein"
author: "Martin Chan"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to textworks: Text Wrangling with Wittgenstein}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
# Load the textworks package
# For development: devtools::load_all()
library(textworks)
```

## Introduction

The **textworks** package provides a collection of convenient functions for text wrangling, cleaning, and processing. These tools are particularly useful for preparing text data for visualization and reporting. 

In this vignette, we'll explore the main functions using quotes from the Austrian philosopher Ludwig Wittgenstein (1889-1951) as our example text. Wittgenstein's profound insights into language and logic make for perfect examples when demonstrating text manipulation techniques.

## Text Analysis Functions

### Counting Words

Let's start with one of Wittgenstein's most famous quotes about the limits of language:

```{r word-counting}
# A famous Wittgenstein quote
quote1 <- "The limits of my language mean the limits of my world."

# Count the number of words
nwords(quote1)

# Compare with a longer quote about language games
quote2 <- "Language is a labyrinth of paths. You approach from one side and know your way about; you approach the same place from another side and no longer know your way about."

nwords(quote2)

# Count words including pseudo-words (numbers, special characters)
mixed_text <- "In Tractus 6.54, Wittgenstein wrote about the ladder metaphor."
nwords(mixed_text)  # Only alphabetic words
nwords(mixed_text, pseudo = TRUE)  # Including numbers and symbols
```

## Text Formatting Functions

### Capitalizing Text

Wittgenstein's aphorisms often need proper formatting for quotes:

```{r capitalization}
# A quote that needs proper capitalization
quote3 <- "whereof one cannot speak, thereof one must be silent."

# Capitalize the first letter of each word
cap_first(quote3)

# Convert to proper sentence case
sentence_case(quote3)

# Compare the difference
original <- "death is not an event in life: we do not live to experience death."
cat("Original:", original, "\n")
cat("Title Case:", cap_first(original), "\n")
cat("Sentence Case:", sentence_case(original), "\n")
```

### Working with CamelCase

Philosophy texts often contain technical terms that need cleaning:

```{r camelcase}
# Philosophical terms in CamelCase format
terms <- c("LanguageGame", "FormOfLife", "PictureTheory", "LogicalPositivism")

# Convert to readable format
camel_clean(terms)

# Useful for creating readable labels
philosophical_concepts <- "Wittgenstein explored LanguageGames and FormsOfLife in his later work."
camel_clean(philosophical_concepts)
```

## String Manipulation Functions

### Extracting Parts of Strings

Sometimes we need to work with parts of longer philosophical texts:

```{r string-extraction}
# A longer Wittgenstein quote
quote4 <- "Philosophy is a battle against the bewitchment of our intelligence by means of language."

# Get the first 10 characters
str_left(quote4, 10)

# Get the last 12 characters  
str_right(quote4, 12)

# Extract different portions for analysis
str_left(quote4, 20)  # "Philosophy is a batt"
str_right(quote4, 20)  # "means of language."
```

### String Replacement

Wittgenstein's works were published in different editions with varying punctuation:

```{r string-replacement}
# Original quote with repeated word
quote5 <- "Language language language is the source of all philosophical problems."

# Replace the 2nd occurrence of "language"
str_replace_nth(quote5, "language", "itself", n = 2)

# Replace the 3rd occurrence 
str_replace_nth(quote5, "language", "communication", n = 3)
```

### Converting Underscores and Adding Wrappers

Working with file names and formatted text:

```{r underscore-and-wrapping}
# File names from Wittgenstein's works
file_names <- c("tractus_logico_philosophicus", "philosophical_investigations", "blue_and_brown_books")

# Convert underscores to spaces
us_to_space(file_names)

# Add quotation marks around philosophical concepts
concepts <- c("language game", "form of life", "private language")
wrap(concepts)

# Use custom wrappers for emphasis
wrap("The world is everything that is the case", wrapper = "**")
```

## Advanced Text Processing

### Adding Line Breaks

For displaying long quotes in visualizations:

```{r line-breaks}
# A lengthy Wittgenstein quote
long_quote <- "A picture held us captive. And we could not get outside it, for it lay in our language and language seemed to repeat it to us inexorably."

# Break after every 3 words for narrow displays
add_line_breaks(long_quote, nword = 3)

# Break after approximately every 25 characters
add_line_breaks(long_quote, nchar = 25)
```

### Working with N-grams

Analyze the structure of Wittgenstein's language:

```{r ngrams}
# Analyze word patterns in a famous passage
passage <- "What we cannot speak about we must pass over in silence"

# Generate and count bigrams (2-word combinations)
count_ngram(passage, n = 2)

# Generate trigrams (3-word combinations)  
count_ngram(passage, n = 3)

# Look at single words with custom column name
count_ngram(passage, n = 1, count_col = "frequency")
```

## Utility Functions

### Number Formatting

When referencing philosophical works and dates:

```{r number-formatting}
# Publication dates and page numbers
tractus_year <- 1921
investigations_pages <- 1953

# Format with commas for readability
comma(tractus_year)  # Though not needed for 4 digits
comma(investigations_pages)

# More relevant for larger numbers
words_in_investigations <- 89000
comma(words_in_investigations)
```

### Color Conversion for Visualizations

Converting RGB values to hex codes for themed visualizations:

```{r color-conversion}
# Create a philosophical color palette
# Deep blue for logic
logic_blue <- rgb2hex(25, 25, 112)

# Warm brown for earthly concerns  
earthly_brown <- rgb2hex(139, 69, 19)

# Show the hex codes
logic_blue
earthly_brown
```

## Practical Example: Processing a Complete Quote

Let's put it all together with a complete workflow:

```{r complete-example}
# Raw text that needs processing
raw_quote <- "the_philosopher_treats_a_question_like_an_illness"

# Step 1: Convert underscores to spaces
step1 <- us_to_space(raw_quote)
cat("Step 1 - Remove underscores:", step1, "\n")

# Step 2: Capitalize properly
step2 <- cap_first(step1)
cat("Step 2 - Capitalize:", step2, "\n")

# Step 3: Add quotation marks
step3 <- wrap(step2)
cat("Step 3 - Add quotes:", step3, "\n")

# Step 4: Count words for metadata
word_count <- nwords(step1)
cat("Word count:", word_count, "\n")

# Create a formatted display
cat("\nFormatted quote:")
cat("\n", step3)
cat("\n- Ludwig Wittgenstein")
cat("\n(", word_count, "words)\n")
```

## Working with Multiple Quotes

Finally, let's process a collection of Wittgenstein quotes:

```{r multiple-quotes}
# Collection of quotes in various formats
quotes_raw <- c(
  "the_world_is_the_totality_of_facts",
  "whereof one cannot speak thereof one must be silent", 
  "languageGames are forms of life",
  "a picture is a model of reality"
)

# Process all quotes
quotes_processed <- quotes_raw %>%
  us_to_space() %>%          # Remove underscores
  camel_clean() %>%          # Fix CamelCase  
  sentence_case() %>%        # Proper sentence capitalization
  wrap()                     # Add quotation marks

# Display the results
for(i in 1:length(quotes_processed)) {
  cat(quotes_processed[i], "\n")
  cat("(", nwords(quotes_raw[i]), "words)\n\n")
}
```

## Conclusion

The **textworks** package provides a comprehensive toolkit for text manipulation in R. Whether you're preparing text for visualization, cleaning data for analysis, or formatting content for reports, these functions offer convenient and intuitive solutions.

As Wittgenstein might have said about good tools: "The tool does not draw attention to itself, but enables the work to be done." The textworks functions are designed to be simple, reliable, and focused on getting your text processing tasks completed efficiently.

For more information about specific functions, use `?function_name` or explore the package documentation.
