---
title: "textworks Function Reference"
author: "Martin Chan"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{textworks Function Reference}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
# Load the textworks package
library(textworks)
```

## Function Reference

This vignette provides a quick reference for all textworks functions with examples.

### Text Analysis Functions

| Function | Purpose | Example |
|----------|---------|---------|
| `nwords()` | Count words in text | `nwords("hello world")` → `2` |

```{r nwords-demo}
# Basic word counting
nwords("The quick brown fox")
nwords("Test@email.com", pseudo = TRUE)  # Include non-alphabetic
```

### Text Formatting Functions

| Function | Purpose | Example |
|----------|---------|---------|
| `cap_first()` | Capitalize first letters | `cap_first("hello world")` → `"Hello World"` |
| `sentence_case()` | Convert to sentence case | `sentence_case("hello. world")` → `"Hello. World"` |

```{r formatting-demo}
cap_first("text wrangling tools")
sentence_case("this is great. really useful.")
```

### String Manipulation Functions

| Function | Purpose | Example |
|----------|---------|---------|
| `str_left()` | Extract from left | `str_left("hello", 3)` → `"hel"` |
| `str_right()` | Extract from right | `str_right("hello", 3)` → `"llo"` |
| `str_replace_nth()` | Replace nth occurrence | `str_replace_nth("test test", "test", "demo", 2)` |

```{r string-demo}
str_left("textworks", 4)
str_right("textworks", 5)
str_replace_nth("foo foo foo", "foo", "bar", n = 2)
```

### Text Cleaning Functions

| Function | Purpose | Example |
|----------|---------|---------|
| `camel_clean()` | Fix CamelCase | `camel_clean("CamelCase")` → `"Camel Case"` |
| `us_to_space()` | Replace underscores | `us_to_space("hello_world")` → `"hello world"` |
| `wrap()` | Add wrapper characters | `wrap("text", "*")` → `"*text*"` |

```{r cleaning-demo}
camel_clean("MyVariableName")
us_to_space("file_name_example")
wrap("important", "**")
```

### Advanced Functions

| Function | Purpose | Example |
|----------|---------|---------|
| `add_line_breaks()` | Add line breaks | `add_line_breaks("long text here", nword = 2)` |
| `count_ngram()` | Count n-grams | `count_ngram("hello world hello", n = 2)` |
| `comma()` | Format numbers | `comma(1000)` → `"1,000"` |
| `rgb2hex()` | Convert RGB to hex | `rgb2hex(255, 0, 0)` → `"#FF0000"` |

```{r advanced-demo}
add_line_breaks("This is a long sentence that needs breaks", nword = 3)
count_ngram("the quick brown fox", n = 2)
comma(1234567)
rgb2hex(100, 150, 200)
```

## Quick Examples by Use Case

### Data Cleaning Pipeline
```{r pipeline}
# Clean messy variable names
messy_names <- c("customer_Name", "totalRevenue", "satisfaction_Score")
clean_names <- messy_names %>%
  us_to_space() %>%
  camel_clean() %>%
  cap_first()

data.frame(Original = messy_names, Cleaned = clean_names)
```

### Text Analysis
```{r analysis}
text <- "R is great for data analysis and visualization"
cat("Words:", nwords(text))
cat("\nBigrams:\n")
print(count_ngram(text, n = 2))
```

### Formatting for Display
```{r display}
value <- 1234567
formatted <- comma(value)
wrapped <- wrap(formatted, "$")
cat("Formatted currency:", wrapped)
```
